New numerical integration of orbits of irregular satellites and prediction of stellar occultations up to 2020

### 2016-03-10 Irregular-occ/Irregular-occ.tex
.

For Phoebe, we updated the ephemeris of \cite{Desmars2013} using the observations of \cite{GomesJunior2015}, observations from Minor Planet Center and observations from Flagstaff. For Nereid and Triton we used the most recent ephemeris available.

Also, with new and precise ephemeris, we present stellar occultation predictions for these 11 objects up to 2020. Phoebe, being the most studied object with a good measured size, can be used to calibrate and evaluate the technique for similar objects.

Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but still studied together by many authors \citep{Sheppard2005, Jewitt2007}. Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007, Elliot_2000}. For all these reasons, Triton was also included as a target in this work.

Excluding Triton \citep{Olkin1997, Elliot_2000}, no observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have discouraged earlier tries. But, in fact, given their relatively closer distances as compared to TNOs and Centaurs, and considering the current precision of their ephemeris and of star positions, we can now reliably predict the exact location and instant where the shadow of the occultation will cross the Earth. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, if the accumulated error (2 sigma or 95\% confidence level) of ephemeris and star position be around 70 mas, we have a probability of about 30\% of observing a stellar occultation, which is quite satisfactory today (see discussion in section \ref{Sec: discussion}).

@@

@@

For Triton, we used the most recent ephemeris published by \cite{Emelyanov2015}. In Fig. \ref{Fig:eph-Triton} we compare it with the ephemeris published in \cite{Zhang2014} and the NEP081 JPL ephemeris \citep{Jacobson2009}. The offsets between the three ephemeris are smaller than 15 mas for the period 2015-2018. This value is much smaller than the apparent size of Triton (124 mas, see Table \ref{Tab: satellite-diameter}), which indicates a good probability of observing an event by this object.

>>>>

For Phoebe, we updated the ephemeris of \cite{Desmars2013} using the observations of \cite{GomesJunior2015}, observations from Minor Planet Center and observations from Flagstaff. %For Nereid and Triton we used the most recent ephemeris available.

Also, with new and precise ephemeris, we present stellar occultation predictions for these 11 objects up to 2020. Phoebe, being the most studied object with a good measured size, can be used to calibrate and evaluate the technique for similar objects. Triton, although it is not an actual irregular satellite, was also include in the list because of its atmosphere which has a complicated and extreme seasonal cycle \citep{McKinnon2007, Elliot_2000}.

%Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but still studied together by many authors \citep{Sheppard2005, Jewitt2007}. Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007, Elliot_2000}. For all these reasons, Triton was also included as a target in this work.

Excluding Triton \citep{Olkin1997, Elliot_2000}, no observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have discouraged earlier tries. But, in fact, given their relatively closer distances as compared to TNOs and Centaurs, and considering the current precision of their ephemeris and of star positions, we can now reliably predict the exact location and instant where the shadow of the occultation will cross the Earth. %For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, if the accumulated error (2 sigma or 95\% confidence level) of ephemeris and star position be around 70 mas, we have a probability of about 30\% of observing a stellar occultation, which is quite satisfactory today (see discussion in section \ref{Sec: discussion}).

@@

%Since 2009 many successful observations of stellar occultations by TNOs have been reported in the literature \citep{Elliot2010, Sicardy2011, Ortiz2012, Braga-Ribas2013}, the main disadvantages in their prediction being large heliocentric distances and ephemeris error, facts somewhat compensated for the larger diameters involved. In contrast to TNOs, the irregular satellites have much better ephemeris because the orbits of their host planets are better  known, their observational time span is much wider and covers many orbital periods. Moreover, the irregular satellites are much closer to Earth which implies in a much smaller shadow path error in kilometers. These advantages may be somewhat balanced by the smaller sizes estimated for the irregular satellites.Thus, in comparison, the chances for a successful observation of an stellar occultation by an irregular satellite should be considered at least also as good as those by TNOs.

@@

Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but still studied together by many authors \citep{Sheppard2005, Jewitt2007}. Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere.

For this satellite, we used the most recent ephemeris published by \cite{Emelyanov2015} which is based on more observations than previous ephemeris and covers 4 years more than the ephemeris of \cite{Jacobson2009}. In Fig. \ref{Fig:eph-Triton} we compare it with the ephemeris published in \cite{Zhang2014} and the NEP081 JPL ephemeris \citep{Jacobson2009}. The offsets between the three ephemeris are smaller than 15 mas for the period 2015-2018. This value is much smaller than the apparent size of Triton (124 mas, see Table \ref{Tab: satellite-diameter}), which indicates a good probability of observing an event by this object.
